---
title: "Salaries"
author: "Mr. Garcia, Mr. Kalyan"
date: "January 22, 2016"
output: html_document
---

##Abstract
The objective of this project is to perform a descriptive analyses of the San Francisco's salaries dataset in order to find if there exists links betweeen the variables of this dataset and answer different questions. It is required to explore, and clean the data in order to be able further analyse it using SPAD software.

##Intro 
In order for San Francisco to become a more transparent city, they decided to release a dataset of the Salaries of the city from the years 2010-2014. We chose this [dataset](http://transparentcalifornia.com/salaries/san-francisco/) in order to answer the following questions: 

1. Is there any links or correlation between the Salaries of the residents of San Francisco?

2. Is there a link between Job Title and the working Status (Full-Time, Part-Time)?

3. Can we use Salary to explain the Job Title or the working Status?

In order to pursue this study, discriminant analysis will be applied, utilizing the methods "PCA, CA and FDA".

##Dataset
###Loading
```{r, message=FALSE}
library(dplyr)
data<-read.csv("Salaries.csv")
```
###Exploring
```{r,}
dim(data)
```
This dataset contains `r dim(data)[2]` variables and `r dim(data)[1]` observations. Which we know in advance that it is more than the amount we can read in SPAD. For this reason the data will be futher explore in order to determine what are the important variables for our study.

###Description
The dataset has the following variables:

| Name | Class | Description |
| ------------ | ------------ | ------------ |
|  `r names(data)[1]` | Int | Id of the dataframe | 
|  `r names(data)[2]` | Factor | Name of the employee |
|  `r names(data)[3]` | Factor | Title of the job position  | 
|  `r names(data)[4]` | Factor | Base anual salary | 
|  `r names(data)[5]` | Factor | Total Overtime payment | 
|  `r names(data)[6]` | Factor | Other payments | 
|  `r names(data)[7]` | Factor | Anual extra benefits received | 
|  `r names(data)[8]` | Numerical | Total salariy without Benefits | 
|  `r names(data)[9]` | Numerical | Total Payment including Benefits | 
|  `r names(data)[10]` | Int | Year (2010-2014) | 
|  `r names(data)[11]` | Logical | Notes are Empty | 
|  `r names(data)[12]` | Factor | Place (San Francisco for all the observations) | 
|  `r names(data)[13]` | Factor | Status of the Employee (Full Time, Part Time) | 

```{r,}
str(data)
summary(data)
```
Reducing the dataset:

1. Remove variables: `Id, EmployeeName, Notes , Agency` will not be useful for the analysis, so they will be removed.

2. Filter the year: The dataset contains 3 years of data, for the purpose of this study, the focus will rely on year 2014.

3. Filter variable "TotalPayBenefits": the values below cero are going to be considered as a mistake and they will be removed.

4. Remove values: The empty values for "Status" will be removed

5. Classes: `BasePay, Overtime Pay, Other Pay`and `Benefits` will be turned into numericals.

###Preprocessing
```{r,}
nums<- c("BasePay","OvertimePay","OtherPay","Benefits")
stat<- c("FT","PT")
dataclean<-data %>% 
            select(-Id, -EmployeeName, -Notes, -Agency) %>% 
            filter(Year==2014, TotalPayBenefits>=0, Status == stat ) #%>% 
            #mutate_each(funs(as.numeric),nums)
dataclean[,nums]<- lapply(dataclean[,nums], as.numeric)
dim(dataclean)
str(dataclean)
summary(dataclean)
```

At this point, the dataframe was reduced to: `r dim(dataclean)[2]` variables and `r dim(dataclean)[1]` observations, out of which 7 are numerical and 2 categorical.

The variable `JobTitle` (categorical), is hard to analyse by having over 800 different categories, and since SPAD will reduce the dataset, a decision was made to filter the dataset by the top 10 most popular `Job Titles` in San Francisco.

```{r}
#Set the categories and order them
categories<-dataclean %>% group_by(JobTitle) %>% summarise(n=n()) 
ordered<-categories[order(-categories$n),]
OrderedCategories<-ordered %>% mutate(cumsum= cumsum(n))
Cat<-as.vector(unlist(OrderedCategories[1:10,1]))
```

```{r}
dataclean1<- dataclean %>% filter(JobTitle==Cat[1])
dataclean2<- dataclean %>% filter(JobTitle==Cat[2])
dataclean3<- dataclean %>% filter(JobTitle==Cat[3])
dataclean4<- dataclean %>% filter(JobTitle==Cat[4])
dataclean5<- dataclean %>% filter(JobTitle==Cat[5])
dataclean6<- dataclean %>% filter(JobTitle==Cat[6])
dataclean7<- dataclean %>% filter(JobTitle==Cat[7])
dataclean8<- dataclean %>% filter(JobTitle==Cat[8])
dataclean9<- dataclean %>% filter(JobTitle==Cat[9])
dataclean10<-dataclean %>% filter(JobTitle==Cat[10])
datafinal<-rbind(dataclean1,dataclean2,dataclean3,dataclean4,dataclean5,
                 dataclean6,dataclean7,dataclean8,dataclean9,dataclean10)
dim(datafinal)
```
###Retrieve dataset
```{r,}
write.table(datafinal,"salariesclean.txt", quote= FALSE, sep ="\t", row.names=FALSE)
```

Now this is the dataset which will be uploaded to SPAD in order to start the analysis.
